---
output:
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# iformr

## Overview

This package provides a collection of tools to leverage the power of the iFormBuilder API. The package is in a developmental state. Additional functions will be added as time allows. Contributions are always welcome. A big thanks to Bill DeVoe of the Maine Department of Marine Resources for several contributions. Please see function documentation for author credits. 

iFormBuilder <https://www.iformbuilder.com/> provides a commercial platform for mobile data collection. The platform includes a number of web-based tools to build and distribute mobile forms, assign users, edit data, and retrieve data from the cloud. Unfortunately, the point-and-click web interface can become a major bottleneck as data volume increases, or you try to push the limits of form design.

Certain tasks, such a preloading large option-lists with newly collected data, are often impractical unless they can be automated. Fortunately, the platform also includes an excellent *Application Programming Interface* (API) http://docs.iformbuilder.apiary.io/#.

Just about anything that can be done using the iFormBuilder web interface can be done infinitely faster using the API. However, for a majority of users, requesting access tokens, or learning the correct syntax to navigate the API can be an insurmountable hurdle. The functions in this repository aim to lower the threshold so that biologists, or field staff with only limited R programming experience will feel comfortable working with the API. The aim is to make form building, data collection, and data retrieval more efficient. 

This package was inspired by the Hadley Wickham's httr package. In particular the *oauth-server-side.R* file at: https://github.com/r-lib/httr/blob/master/R/oauth-server-side.R. Encoding and signing JSON web tokens now relies on the `jose` package by Jeroen Ooms https://github.com/jeroen/jose. Thanks to Danny Warren (WDFW) for identifying issues where requests for access tokens would occasionaly fail. That issue should now be fixed by using the `jose` package to sign and encode JSON web tokens.

## Installation

You will need to install R-3.4.1 or higher.

```{r, eval = FALSE}
# Install the development version from GitHub:
# install.packages("remotes")
remotes::install_github("arestrom/iformr")
```

## Storing API connection credentials

In order to use the functions in this package you **must** first get a `client_key` and `client_secret` for your profile from iFormBuilder. If you have a dedicated server this can be done by logging in to your iFormBuilder server as a *server admin* and clicking on the *Server Admin* tab. Select *Manage Profiles*, highlight the row for the desired profile then click *Manage*. Next select *Server Admin* > *API Apps* > *New Client* and follow the directions to create a new client. You will be asked to assign the API application to an existing *Username*. It is good practice to create a separate, dedicated *user* for any API requests. After creating the new client, a new *Client Key* and *Client Secret* will be posted to the *API Applications* table. 

If you do not have dedicated server you will need to contact iFormBuilder, or check their webpage, for directions on how to obtain a `client_key` and `client_secret`. 

To add the `client_key` and `client_secret` to your `.Renviron` file you will first need to create, or locate, your `.Renviron` file. In your R console, enter: 

```{r, eval = FALSE}
normalizePath("~/")
```

The path to your `.Renviron` file will be displayed. Use RStudio to either create a new `.Renviron` file if it does not exist, or open your existing `.Renviron` file from the specified location. The `.Renviron` is a simple text file and will look similar to:

```
MyProfileNameClientKey=eb7411c1525c486ae7a64727965c3e042eab7389
MyProfileNameClientSecret=471dbbe5061f4c1db9141c48cc1d0d5a96e60cc6

```

Simply add your new `client_key` and `client_secret` as in the example above. The *client_key_name* is the section before the equals sign. The `client_key` is the value following the equals sign. The same pattern is used for the `client secret`. You can enter any convenient name for the name portion. These names are what you will enter as arguments to the `get_iform_access_token()` function to request an `access_token`.

Make sure there is an extra space at the bottom of the last entry in the `.Renviron` file before saving, or R will silently fail to read the entries. You will need to close RStudio and then reopen after saving. Afterwards, every time you open RStudio the environment variables will be available in your session.

Detailed instructions on how to add credentials to your `.Renviron` file can be found in the Appendix section of: https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html

Obtaining an `access_token` is a prerequisite for any requests made to the iFormBuilder API.

## Request limits and usage notes

Please see the introduction section of the iFormBuilder [API](http://docs.iformbuilder.apiary.io/#reference/user-resource/user) for general notes on API usage and limits. 

There are limits to the number of items you can retrieve per request. You can typically request up to `1000` records of data from individual forms or subforms. The same `1000` record limit applies to requests for items in option lists. Other request are limited to `100` items. For example, if you need to retrieve a list of the names of all option lists in your profile, you can only retrieve `100` in each request. This is more fully explained in the link above. If you hit these limits you can simply make additional requests and set the `offset` parameter to skip previously returned records.

The introduction section provides examples of how to specify the *fields* you need to return from any given request, and how to set request *parameters*. **Please read** the [API introduction](http://docs.iformbuilder.apiary.io/#reference/user-resource/user) before proceeding.

## Building mobile forms

Contributions by Bill DeVoe (William.DeVoe@maine.gov) provide a set of tools to generate the basic structure of mobile forms directly from R dataframes. See the `create_page()`, `copy_page()`, `rename_page()`, and `data2form()` functions. By programatically generating mobile forms from data, the otherwise tedious task of creating mobile forms by hand can be standardized and made considerably more reliable and efficient. 

For those just beginning to learn form design, [Mobile forms for salmon surveys](https://arestrom.github.io/mf4ss/) may serve as a useful starting point. This online text steps through the manual process of form design using the standard iFormBuilder toolset. 






