---
output:
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# iformr

## Overview

This package provides a collection of tools to leverage the power of the iFormBuilder API. The package is in an early-development state. Additional functions will be added as time allows.

iFormBuilder <https://www.iformbuilder.com/> provides a commercial platform for mobile data collection. The platform includes a number of web-based tools to build and distribute mobile forms, assign users, edit data, and retrieve data from the cloud. Unfortunately, the point-and-click web interface can become a major bottleneck as data volume increases, or you try to push the limits of form design.

Certain tasks, such a preloading large option-lists with newly collected data, are often impractical unless they can be automated. Fortunately, the platform also includes an excellent *Application Programming Interface* (API) http://docs.iformbuilder.apiary.io/#.

Just about anything that can be done using the iFormBuilder web interface can be done infinitely faster using the API. However, for a majority of users, requesting access tokens, or learning the correct syntax to navigate the API can be an insurmountable hurdle. The functions in this repository aim to lower the threshold so that biologists, or field staff with only limited R programming experience will feel comfortable working with the API. The aim is to make form building, data collection, and data retrieval more efficient. 

This package was inspired by the Hadley Wickham's httr package. In particular the *oauth-server-side.R* file at: https://github.com/r-lib/httr/blob/master/R/oauth-server-side.R. The *hex_to_raw()* function has been imported from the openssl package by Jeroen Ooms. Future plans are to shift to the *jose* package https://github.com/jeroen/jose for signing and verifying JSON web tokens.

## Installation

There was a bug in R-3.4.0 that prevented installation of this package. You will need to install R-3.4.1 or higher.

```{r, eval = FALSE}
# Install the development version from GitHub:
# install.packages("devtools")
devtools::install_github("arestrom/iformr")
```

## Storing API connection credentials

In order to use the functions in this package you **must** first get a `client_key` and `client_secret` for your profile from iFormBuilder. If you have a dedicated server this can be done by logging in to your iFormBuilder server as a *server admin* and clicking on the *Server Admin* tab. Select *Manage Profiles*, highlight the row for the desired profile then click *Manage*. Next select *Server Admin* > *API Apps* > *New Client* and follow the directions to create a new client. You will be asked to assign the API application to an existing *Username*. It is good practice to create a separate, dedicated *user* for any API requests. After creating the new client, a new *Client Key* and *Client Secret* will be posted to the *API Applications* table. 

If you do not have dedicated server you will need to contact iFormBuilder, or check their webpage, for directions on how to obtain a client_key and client_secret. 

To add the client_key and client_secret to your `.Renviron` file you will first need to create, or locate, your `.Renviron` file. In your R console, enter: 

```{r, eval = FALSE}
normalizePath("~/")
```

The path to your `.Renviron` file will be displayed. Use RStudio to either create a new `.Renviron` file if it does not exist or open the existing `.Renviron` file from the specified location. The `.Renviron` is a simple text file and will look similar to:

```
MyProfileNameClientKey=eb7411c1525c486ae7a64727965c3e042eab7389
MyProfileNameClientSecret=471dbbe5061f4c1db9141c48cc1d0d5a96e60cc6

```

Simply add your new `client_key` and `client_secret` as in the example above. The *client_key_name* is the section before the equals sign. The `client_key` is the value following the equals sign. The same pattern is used for the `client secret`. You can enter any convenient name for the name portion. These names are what you will enter as arguments to the `get_iform_access_token()` function to request an `access_token`.

Make sure there is an extra space at the bottom of the last entry in the `.Renviron` file before saving, or R will silently fail to read the entries. You will need to close RStudio and then reopen after saving. Afterwards, every time you open RStudio the environment variables will be available in your session.

Detailed instructions on how to add credentials to your `.Renviron` file can be found in the Appendix section of: https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html

Obtaining an `access_token` is a prerequisite for any requests made to the iFormBuilder API.








